---
title: "First-pass refinement of putative NK signature genes"
author: "C Shembrey"
date: "Created 20/4/2020, Cleaned 21/06/2022"
output: html_document
---

---
In this document, we refine our initial collated NK signature (295 genes taken from the literature, previously published signatures, and our in-house DE analysis).

NOTE: filtration of the gene signature derived at the end of this document is continued (further refinements) in the "Tumour and Stromal filt of post-scRNAseq 82 gene sig.Rmd" file

# Setup
```{r results = 'hide', message = FALSE, warning = FALSE}
library(limma)
library(scales)
library(ggplot2)
library(ggrepel)
library(plotly)
library(gridExtra)
library(grid)
library(ggpubr) # for statistics

source("../Files/biplot_ggplot.R")
```

# NK Signature Filtration using Biplot

## Import NK Signatures to Test:
See "Collating NK Signature Sources - version control.Rmd" for details re: how this geneset was compiled
```{r}
nk <- read.csv("../Files/Initial_NK_Sig_295_Genes.csv") 
rownames(nk) <- nk$GeneID
```

# Generate table of NK Signature sources:
```{r}
CiberAct <- read.csv("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/data/NK_Signatures/Compiled/Cibersort_NK-Active.csv")
CiberAct <- nk$Symbol %in% CiberAct$Genes
SumTable <-  cbind(nk, CiberAct)

CiberRest <- read.csv("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/data/NK_Signatures/Compiled/Cibersort_NK-Resting.csv")
Cibersort_Resting <- nk$Symbol %in% CiberRest$Genes
SumTable <-  cbind(SumTable, Cibersort_Resting)

CursonsSups <- read.csv("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/data/NK_Signatures/Compiled/CursonsSups_112Genes.csv")
CursonsSups <- nk$Symbol %in% CursonsSups$Cursons.Supplementary.Data
SumTable <-  cbind(SumTable, CursonsSups)

# JerbyMarker <- read.csv("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/data/NK_Signatures/Compiled/Jerby_NK.Marker.csv")
# JerbyMarker <- nk$Symbol %in% JerbyMarker$Symbol
# SumTable <-  cbind(SumTable, JerbyMarker)
# 
# Jerby_scDEG <- read.csv("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/data/NK_Signatures/Compiled/Jerby_scDEGs.csv")
# Jerby_scDEG <- nk$Symbol %in% Jerby_scDEG$Symbol
# SumTable <-  cbind(SumTable, Jerby_scDEG)

MergedDEGs_Bulk <- readRDS("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/data/KeyData/blood_nk/NK_genelist_2bld.RDS")
NK_upSet <- MergedDEGs_Bulk$NK_upSet
DEGs_2BldBulk <- nk$Symbol %in% NK_upSet
SumTable <-  cbind(SumTable, DEGs_2BldBulk)

DEGs_NKvsPanT <- MergedDEGs_Bulk$NK_panT_up
DEGs_NKvsPanT <- nk$Symbol %in% DEGs_NKvsPanT
SumTable <-  cbind(SumTable, DEGs_NKvsPanT)

RecepAct <- read.csv("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/data/NK_Signatures/Compiled/NK_ReceptorActivating.csv")
RecepInhib <- read.csv("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/data/NK_Signatures/Compiled/NK_ReceptorInhibitory.csv")
Receptors <- rbind(RecepAct, RecepInhib) # remove dual function receptors
Receptors <- nk$Symbol %in% Receptors$Genes
SumTable <-  cbind(SumTable, Receptors)

WangGenes <- read.csv("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/data/GeneSignatures/Immune_stroma/Stroma_Immune_markers_Wang_2017_Gastroenterology.csv", stringsAsFactors = F)
WangGenes <- WangGenes[WangGenes$Marker.type == "NK cell", ]
WangGenes <- WangGenes$Gene.Name
WangGenes <- nk$Symbol %in% WangGenes
SumTable <-  cbind(SumTable, WangGenes)

#write.csv(outPath, "200501_NKSig_303GeneProvenance.csv")
```

# How many genes have multiple sources?
```{r}
colnames(SumTable)
TstHtmp <- SumTable[ , c(3,6:12)]
rownames(TstHtmp) <- TstHtmp$Symbol
TstHtmp <- TstHtmp[ , -c(1)]

TstHtmp$SumSource <- rowSums(TstHtmp)
table(TstHtmp$SumSource)

which(TstHtmp$SumSource == 7) # zero genes in all 7 sources
which(TstHtmp$SumSource == 6) # KLRD1, KIR3DL2
which(TstHtmp$SumSource == 5) # CD244, KIR2DS4, KLRC3, KLRF1, KLRK1, KIR2DL1, KLRB1, IL12RB

#pdf("Histogram_GeneSigSumOfSources.pdf")
ggplot(TstHtmp, aes(SumSource)) +
       geom_bar(width = 1, fill = "forestgreen", color = "black") +
       geom_text(stat = "count", color = "forestgreen", aes(label = ..count..), vjust=-1) +
  scale_x_continuous(name = "Sum of Sources", breaks = c(1,2,3,4,5,6,7)) +
  scale_y_continuous(name = "Count (n of Genes)", breaks = c(0,50,100,150,200, 250)) +
  ylim(c(0,200)) + 
  labs(x = "Sum of Sources", y = "Count (n of Genes)") +
  theme_light() +
  theme(panel.grid.major = element_blank())
dev.off()

# add SumSource info to nk dataframe:
all(nk$Symbol == rownames(TstHtmp)) # T
nk$SumSource <- TstHtmp$SumSource
```

## Import and prepare the log-normalised data for the Linsley bulkRNAseq blood data:
```{r}
dataList <- readRDS("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/Cleaned_Code_For_Gitlab/Files/CancerBlood_dataList_CRC_logRPKM.RDS") 
LinsBld <- dataList$bldCells_logRPKM

# subset to NK sig
Bld_NKSig <- LinsBld[rownames(LinsBld) %in% nk$GeneID, ] # 390 genes x 28 samples --> there are 5 missing genes (KIR2DS3, KIR2DL2, KIR2DL5B, TRBC1, MIR181A2HG)

# split the annotation file into its constituent cell types:
LinsAnnot <- dataList$bldCells_annot 
LinsAnnot$celltype_s <- as.character(LinsAnnot$celltype_s)

annotSplit <- split(LinsAnnot, LinsAnnot$celltype_s, drop = F) # all elements (data frames from individual cell types) are 4 samples x 5

NKdata_Lins <- Bld_NKSig[, colnames(Bld_NKSig) %in% rownames(annotSplit$NK)]
B.Cell <- Bld_NKSig[, colnames(Bld_NKSig) %in% rownames(annotSplit$`B-cells`)]
CD4 <- Bld_NKSig[, colnames(Bld_NKSig) %in% rownames(annotSplit$CD4)]
CD8 <- Bld_NKSig[, colnames(Bld_NKSig) %in% rownames(annotSplit$CD8)]
Mono <- Bld_NKSig[, colnames(Bld_NKSig) %in% rownames(annotSplit$Monocytes)]
Neut <- Bld_NKSig[, colnames(Bld_NKSig) %in% rownames(annotSplit$Neutrophils)]

nonNKs <- Bld_NKSig[, ! colnames(Bld_NKSig) %in% rownames(annotSplit$NK)] # 279 genes x 24 samples (all but NKs)

# generate list for plotting later
CellTypeData_Linsley <- list(Bcell = B.Cell,
                     CD4 = CD4,
                     CD8 = CD8,
                     Monocytes = Mono,
                     Neutrophils = Neut)

ctdNames_Lins <- c("B.Cell",
                 "CD4",
                 "CD8",
                 "Mono",
                 "Neut")

# Update table to track subsetting
Linsley <- nk$GeneID %in% rownames(Bld_NKSig)
SumTable <- cbind(SumTable, Linsley)
```

## Generate biplots for the NK Sig using the Linsley data

```{r fig.width= 10, fig.height = 10}
# Subset NK sig (395 genes) so that only the 298 genes that appear in LinsBld are kept
NK <- nk[rownames(nk) %in% rownames(Bld_NKSig), ]
NK <- NK[rownames(Bld_NKSig), ]

# generate annotations for biplot
sourceAnnot <- NK$SumSource
names(sourceAnnot) <- rownames(NK)

geneAnnot <- NK$Symbol
names(geneAnnot) <- rownames(Bld_NKSig)

p <- list()
#pdf("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/200501_Linsley_biplots_NK-0.5_Blood-0.5_Lax.pdf", width = 25, height = 8)
pdf("Linsley_Biplots.pdf", width = 30, height = 12, useDingbats = F)
for(i in ctdNames_Lins){
  ctd <- get(i)
  #print(
  p[[i]] <- biplot(
  data.expr = NKdata_Lins,
  data.noexpr = ctd,
  data.names = c("NK Cells_Linsley", i),
  data.expr.th = 0.5,
  data.noexpr.th = 0.5,
  annot = sourceAnnot,
  annot.name = "Sum of Sources",
  #point.labels = geneAnnot,
  text.size = 1.2) + 
    geom_vline(xintercept = 0, col = "darkred", linetype = "dashed") +
    #geom_vline(xintercept = -1, col = "darkred", linetype = "dashed") +
   # geom_hline(yintercept = 3, col = "darkred", linetype = "dashed") +
    geom_hline(yintercept = 2, col = "darkred", linetype = "dashed") +
    theme(aspect.ratio = 0.4) # legend.position = "none"
  #)
}

PlotTitle <- textGrob(paste0("Linsley Data: median expression of NK cell signature genes in NK cells vs other immune cell types \n "), gp=gpar(fontface="bold", fontsize=20))

do.call(grid.arrange, c(p, list(ncol = 3, top = PlotTitle)))
dev.off()

# Bind this to the Xu data and add a single legend below (or at right)
```

## Filter genes based on biplot strict thresholds (dashed lines) & lax thresholds (solid lines)
```{r}
data.expr.th <- 0.5
data.noexpr.th <- 0.5

LinsStrict <- names(which(
              apply(NKdata_Lins, 1, quantile, probs = data.expr.th) >= 3 &
              apply(CD8, 1, quantile, probs = data.noexpr.th) <= -1 &
              apply(CD4, 1, quantile, probs = data.noexpr.th) <= -1 &
              apply(B.Cell, 1, quantile, probs = data.noexpr.th) <= -1 &
              apply(Mono, 1, quantile, probs = data.noexpr.th) <= -1 &
              apply(Neut, 1, quantile, probs = data.noexpr.th) <= -1
              ))
length(LinsStrict) # 8 genes (when we use NK expression threshold = 0.25 and blood cell type expression threshold = 0.75, there were 29 genes...)
nkSymbols <- NK[LinsStrict, ]
nkSymbols$Thresh <- "BiplotStrict"
#write.csv(nkSymbols, "191126_NKSig_15Genes_BiplotMedians_StrictThreshold.csv")

LinsLax <- names(which(
              apply(NKdata_Lins, 1, quantile, probs = data.expr.th) >= 2 &
              apply(CD8, 1, quantile, probs = data.noexpr.th) <= 0 &
              apply(CD4, 1, quantile, probs = data.noexpr.th) <= 0 &
              apply(B.Cell, 1, quantile, probs = data.noexpr.th) <= 0 &
              apply(Mono, 1, quantile, probs = data.noexpr.th) <= 0 &
              apply(Neut, 1, quantile, probs = data.noexpr.th) <= 0
              ))

length(LinsLax) # 36 genes (when we use NK expression threshold = 0.25 and blood cell type expression threshold = 0.75, there were 91 genes...)
nkSymbols_lax <- NK[LinsLax, ]
nkSymbols_lax$Thresh <- "BiplotLax"
#write.csv(nkSymbols_lax, "191126_NKSig_36genes_BiplotMedians_LaxThreshold.csv")
length(nkSymbols_lax$GeneID)

overlap <- intersect(nkSymbols$GeneID, nkSymbols_lax$GeneID)
lax_only <- nkSymbols_lax[! nkSymbols_lax$GeneID %in% overlap, ]
lax_only$Thresh <- "BiplotLax" 
ThreshData <- rbind(nkSymbols, lax_only) # 36 genes (8 with "strict" and 28 with "lax" annotations)

#write.csv(ThreshData, "191126_NKSig__LinsBiplot_MedvsMedFiltered_StrictLaxThresholds.csv")
#ThreshData <- read.csv("191126_NKSig__LinsBiplot_MedvsMedFiltered_StrictLaxThresholds.csv", stringsAsFactors = F)
rownames(ThreshData) <- ThreshData$GeneID 

ThreshData$SumSource <- as.factor(ThreshData$SumSource)
sourceAnnot <- ThreshData$SumSource
names(sourceAnnot) <- rownames(ThreshData)

# plot the 36 x passing genes
plot1 <- lapply(CellTypeData_Linsley, function(x) { x[rownames(ThreshData), , drop = F] })
names(plot1) <- ctdNames_Lins

#pdf("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/Filtered_LinsleyGeneBiplots.pdf", width = 25, height = 8)
p <- list()
for(i in ctdNames_Lins){
  d <- plot1[i]
  p[[i]] <- biplot(
  data.expr = NKdata_Lins[rownames(ThreshData), ],
  data.noexpr = d,
  data.names = c("NK Cells_Lins", i),
  data.expr.th = 0.5,
  data.noexpr.th = 0.5,
  annot = sourceAnnot,
  annot.name = "Sum of Sources",
  point.labels = geneAnnot,
  text.size = 0.75) +
    geom_vline(xintercept = 0, col = "darkred", linetype = "dashed") +
    geom_hline(yintercept = 2, col = "darkred", linetype = "dashed") +
    theme(aspect.ratio = 0.4, legend.position = "none")
  }

PlotTitle <- textGrob(paste0("Linsley Data: median expression of filtered NK cell signature genes in NK cells vs other immune cell types \n "), gp=gpar(fontface="bold", fontsize=20))

do.call(grid.arrange, c(p, list(ncol = 3, top = PlotTitle)))
dev.off()

# add to table
LinsleyPass <- nk$GeneID %in% ThreshData$GeneID
SumTable <- cbind(SumTable, LinsleyPass)
```

## Validate the signature in another blood data; import and prepare the log-normalised data for the Xu bulkRNAseq blood data and repeat the biplot analysis:

```{r fig.width=12}
Xu <-  read.table("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/data/KeyData/blood_nk/XuBulk_logTPMFilt.txt", check.names = F) # 114 samples
Xu_NKSig <- Xu[rownames(Xu) %in% nk$GeneID, ] # 288 genes --> we miss 27 genes

# update table
Xutable <- nk$GeneID %in% rownames(Xu)
SumTable <- cbind(SumTable, Xutable)

summary(Xutable)
# annotation file 
XuAnnot <- read.csv("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/data/KeyData/blood_nk/XuBulk_Annot.csv")
rownames(XuAnnot) <- XuAnnot$Samples
XuAnnot <- XuAnnot[ , -c(1:2)]

# make sample names consistent across annot and expr files (the expr data names get corrected when it becomes as dataframe ie. it can't handle "VD2-" and converts it to "VD2..1":
testvector <- as.character(rownames(XuAnnot))
colnames(Xu_NKSig) <- testvector
all(rownames(XuAnnot) == colnames(Xu_NKSig)) # TRUE

# split to constituent cell types
Xu_annotSplit <- split(XuAnnot, XuAnnot$CellType, drop = F)

Xu_NK <- Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$NK)]
Xu_B <- Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$B)]
Xu_Baso <- Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$Basophils)]
Xu_CD4 <- Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$CD4)]
Xu_CD8 <-Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$CD8)]
Xu_MAIT <-Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$MAIT)]
Xu_mDC <-Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$mDC)]
Xu_Mono <- Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$Monocytes)]
Xu_Neut <- Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$Neutrophils)]
Xu_pDC <- Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$pDC)]
Xu_Plas <- Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$Plasmablasts)]
Xu_Prog <-Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$Progenitor)]
Xu_Tfh <-Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$TFH)]
Xu_Th1 <-Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$Th1)]
Xu_Th1_17 <-Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$`Th1/Th17`)]
Xu_Th17 <- Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$Th17)]
Xu_Th2 <- Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$Th2)]
Xu_Treg <-Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$Treg)]
Xu_gamDelneg <-Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$`VD2-`)]
Xu_gamDelpos <-Xu_NKSig[, colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$`VD2+`)]

nonNK_Xu <- Xu_NKSig[, ! colnames(Xu_NKSig) %in% rownames(Xu_annotSplit$NK)]

# merge the plasmacytoid and myeloid DC subsets
Xu_DC = cbind(Xu_pDC, Xu_mDC)
# merge T-cell subsets to crate "pan-T" category
Xu_panT = cbind(Xu_CD4, Xu_CD8, Xu_MAIT, Xu_Tfh, Xu_Th1, Xu_Th17, Xu_Th1_17, Xu_Th2, Xu_Treg, Xu_gamDelneg, Xu_gamDelpos)

# exclude plasma cells and progenitor cells
CellTypeData_Xu <- list(Xu_B = Xu_B,
                        Xu_Baso = Xu_Baso,
                        Xu_DC = Xu_DC,
                        Xu_Mono = Xu_Mono,
                        Xu_Neut = Xu_Neut,
                        Xu_panT = Xu_panT,
                        Xu_CD4 = Xu_CD4,
                        Xu_CD8 = Xu_CD8,
                        Xu_MAIT = Xu_MAIT,
                        Xu_Tfh = Xu_Tfh,
                        Xu_Th1 = Xu_Th1,
                        Xu_Th1_17 = Xu_Th1_17,
                        Xu_Th17 = Xu_Th17,
                        Xu_Th2 = Xu_Th2,
                        Xu_Treg = Xu_Treg,
                        Xu_gamDelneg = Xu_gamDelneg,
                        Xu_gamDelpos = Xu_gamDelpos)

ctdNames_Xu <- c("Xu_B",
                  "Xu_Baso",
                  "Xu_DC",
                  "Xu_Mono",
                  "Xu_Neut",
                  "Xu_panT",
                  "Xu_CD4",
                  "Xu_CD8",
                  "Xu_MAIT",
                  "Xu_Tfh",
                  "Xu_Th1",
                  "Xu_Th1_17",
                  "Xu_Th17",
                  "Xu_Th2",
                  "Xu_Treg",
                  "Xu_gamDelneg",
                  "Xu_gamDelpos")

# generate annotations for biplot
sourceAnnot <- NK$SumSource
names(sourceAnnot) <- rownames(NK)

geneAnnot <- NK$Symbol
names(geneAnnot) <- rownames(NK)

#pdf("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/200501_XuBiplots.pdf", width = 25, height = 24)
pdf("Xu_Biplots.pdf", width =30, height = 24)

p <- list()
for(i in ctdNames_Xu){
  ctd_Xu <- get(i)
  p[[i]] <- biplot(
  data.expr = Xu_NK,
  data.noexpr = ctd_Xu,
  data.names = c("NK Cells_Xu", (i)),
  data.expr.th = 0.5,
  data.noexpr.th = 0.5,
  annot = sourceAnnot,
  annot.name = "Gene source",
  #point.labels = geneAnnot,
  text.size = 1.2
  ) +
  #geom_vline(xintercept = 1, col = "red", linetype = "dashed") +
  #geom_hline(yintercept = 4, col = "red", linetype = "dashed") +
  geom_vline(xintercept = 2, col = "darkred", linetype = "dashed") +
  geom_hline(yintercept = 2, col = "darkred", linetype = "dashed") +
  theme(aspect.ratio = 0.4) # legend.position = "none"
}

PlotTitle <- textGrob(paste0("Xu Data: median expression of NK cell signature genes in NK cells vs other immune cell types \n "), gp=gpar(fontface="bold", fontsize=20))

do.call(grid.arrange, c(p, list(ncol = 3, top = PlotTitle)))
dev.off()
```

## Plot the 23 x passing genes

NB: There are 23 intersecting genes, when the gammadelta subsets are split (ie. when gamDelneg and gamDelpos are not combined). There are 29 genes when the 2 x gamDel subtypes are combined
```{r}
allXu_intersect <-  sapply(ctdNames_Xu, function(x) {
                    ctd_Xu <- get(x)
                    names(which(
                    apply(Xu_NK, 1, quantile, probs = data.expr.th) >= 2 & 
                    apply(ctd_Xu, 1, quantile, probs = data.noexpr.th) <= 2))})

allXu_intersect <- Reduce(intersect, allXu_intersect)                 
length(allXu_intersect) # 23 genes

# again, change to discrete colour scale here - change biplot viridis script value to 0.3:
sourceAnnot <- as.factor(sourceAnnot)
levels(sourceAnnot)

plot2 <- lapply(CellTypeData_Xu, function(x) { x[allXu_intersect, , drop = F]} )
names(plot2) <- ctdNames_Xu

#pdf("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/200501_Filtered_XuBiplots.pdf", width = 25, height = 24)

p <- list()
for(i in ctdNames_Xu){
  d <- plot2[i]
  p[[i]] <- biplot(
  data.expr = Xu_NK[allXu_intersect, ],
  data.noexpr = d,
  data.names = c("NK Cells_Xu", i),
  data.expr.th = 0.5,
  data.noexpr.th = 0.5,
  annot = sourceAnnot,
  annot.name = "Gene source",
  point.labels = geneAnnot,
  text.size = 0.75
  ) +
  geom_vline(xintercept = 2, col = "darkred", linetype = "dashed") +
  geom_hline(yintercept = 2, col = "darkred", linetype = "dashed") +
  theme(aspect.ratio = 0.4, legend.position = "none")
  }

PlotTitle <- textGrob(paste0("Xu Data: median expression of filtered NK cell signature genes in NK cells vs other immune cell types \n "), gp=gpar(fontface="bold", fontsize=20))

#pdf("TestDingbats.pdf", useDingbats = F, width = 20, height = 15)
do.call(grid.arrange, c(p, list(ncol = 3, top = PlotTitle)))
dev.off()

# update table
XuPass <- nk$GeneID %in% allXu_intersect
SumTable <-cbind(SumTable, XuPass)
#write.csv(SumTable, "210123_GeneProvenanceTable_FromRmd1.csv")
# 23/1/21: subset the SumTable to only the NKSigGenes
```

## Plot the genes coming from NK_panT_up only in both data sets. Check that at least some of these genes are conserved when we filter via the biplot methods above, as we need some genes to help differentiate NK cells from T cells.
```{r}
# Xu Data
getIDs <- as.character(NK$GeneID[NK$Source == "NK_panT_up"])  
plot3 <- lapply(CellTypeData_Xu, function(x) { x[getIDs, , drop = F]} )
names(plot3) <- ctdNames_Xu

#pdf("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/200501_Biplots_NKvsPanTGenes_XuData.pdf", width = 25, height = 24)

p <- list()
for(i in ctdNames_Xu){
  d <- plot3[i]
  p[[i]] <- biplot(
  data.expr = Xu_NK[getIDs, ],
  data.noexpr = d,
  data.names = c("NK Cells_Xu", i),
  data.expr.th = 0.5,
  data.noexpr.th = 0.5,
  annot = sourceAnnot,
  annot.name = "Gene source",
  point.labels = geneAnnot,
  text.size = 0.75
  ) +
  geom_vline(xintercept = 2, col = "darkred", linetype = "dashed") +
  geom_hline(yintercept = 2, col = "darkred", linetype = "dashed") +
  theme(aspect.ratio = 0.4, legend.position = "none")
}

PlotTitle <- textGrob(paste0("NK vs T cell subsets: median expression of NK_panT_up signature genes only in the Xu data \n "), gp=gpar(fontface="bold", fontsize=20))

do.call(grid.arrange, c(p, list(ncol = 3, top = PlotTitle)))
dev.off()

sum(getIDs %in% allXu_intersect) # there are 5 genes from the NK_panT_up set in the filtered Xu sig # "1305" (COL13A1)   "494470" (RNF165) "9079" (LDB2)  "79698" (ZMAT4) "158038" (LINGO2)

# Linsley Data
#pdf("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/200501_Biplots_NKvsPanTGenes_LinsleyData.pdf", width = 25, height = 8, useDingbats = F)

plot4 <- lapply(CellTypeData_Linsley, function(x) { x[getIDs, , drop = F]} )
names(plot4) <- ctdNames_Lins

p <- list()
for(i in ctdNames_Lins){
  d <- plot4[i]
  p[[i]] <- biplot(
  data.expr = NKdata_Lins[getIDs, ],
  data.noexpr = d,
  data.names = c("NK Cells_Lins", i),
  data.expr.th = 0.5,
  data.noexpr.th = 0.5,
  annot = sourceAnnot,
  annot.name = "Gene source",
  point.labels = geneAnnot,
  text.size = 0.75
  ) +
  geom_vline(xintercept = 2, col = "darkred", linetype = "dashed") +
  geom_hline(yintercept = 2, col = "darkred", linetype = "dashed") +
     theme(aspect.ratio = 0.4, legend.position = "none")
}

PlotTitle <- textGrob(paste0("NK vs T cell subsets: median expression of NK_panT_up signature genes only in the Linsley data \n "), gp=gpar(fontface="bold", fontsize=20))

do.call(grid.arrange, c(p, list(ncol = 3, top = PlotTitle)))
dev.off()

sum(getIDs %in% ThreshData$GeneID) # there are 3 genes from the NK_panT_up set in the filtered Lins sig: "494470" (RNF165) "9079" (LDB2) "284194 (LGALS9B)"
```

## Take the union of the candidate genes from the Linsley and Xu biplot analyses:

```{r}
allXu_intersect # Xu data - 23 genes (single 2 x 2 threshold)
ThreshData$GeneID # Linsley data - 36 genes (8 from strict thresholds, 28 from lax thresholds)

Cand_NKSig <- union(ThreshData$GeneID, allXu_intersect) # 49 genes
length(intersect(ThreshData$GeneID, allXu_intersect)) # only 10 genes common to both

# add Cand_NKsig to summary table
Cand_NKSig_table <- nk$GeneID %in% Cand_NKSig
SumTable <-  cbind(SumTable, Cand_NKSig_table)

# visualise the 49 candidate genes on both data (change viridis value back to 0.65:
LinsCand <- lapply(CellTypeData_Linsley, function(x) { as.data.frame(x)[Cand_NKSig, , drop = F]})
names(LinsCand) <- ctdNames_Lins

#pdf("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/200501_Biplots_Candidate49Genes_PostLinsleyFiltration.pdf", width = 25, height = 8)

p <- list()
for(i in ctdNames_Lins){
  d <- LinsCand[i]
  p[[i]] <- biplot(
  data.expr = as.data.frame(NKdata_Lins)[Cand_NKSig, ],
  data.noexpr = d,
  data.names = c("NK Cells_Lins", i),
  data.expr.th = 0.5,
  data.noexpr.th = 0.5,
  annot = sourceAnnot,
  annot.name = "Gene source",
  point.labels = geneAnnot,
  text.size = 0.75
  ) + 
  geom_vline(xintercept = 0, col = "darkred", linetype = "dashed") +
  geom_hline(yintercept = 2, col = "darkred", linetype = "dashed") +
  theme(aspect.ratio = 0.4, legend.position = "none")
}

PlotTitle <- textGrob(paste0("Linsley Data: NK vs immune cell subsets: median expression of Cand_NKsig (49 genes) \n "), gp=gpar(fontface="bold", fontsize=20))

do.call(grid.arrange, c(p, list(ncol = 3, top = PlotTitle)))
dev.off()
# it looks great for all cell types!
```

```{r}
# repeat for Xu data
XuCand <- lapply(CellTypeData_Xu, function(x) { x[Cand_NKSig, , drop = F]} )
names(XuCand) <- ctdNames_Xu

#pdf("/Users/shembreycarolyn/Documents/PhD Archive/PhD Project/Computational Biology/200501_Biplots_Candidate49Genes_PostXuFiltration_NoLabels.pdf", width = 25, height = 24)

p <- list()
for(i in ctdNames_Xu){
  d <- XuCand[i]
  p[[i]] <- biplot(
  data.expr = Xu_NK[Cand_NKSig, ],
  data.noexpr = d,
  data.names = c("NK Cells_Xu", i),
  data.expr.th = 0.5,
  data.noexpr.th = 0.5,
  annot = sourceAnnot,
  annot.name = "Gene source",
  point.labels = geneAnnot,
  text.size = 0.75
  ) +
  geom_vline(xintercept = 2, col = "darkred", linetype = "dashed") +
  geom_hline(yintercept = 2, col = "darkred", linetype = "dashed") +
  theme(aspect.ratio = 0.4, legend.position = "none")
}

PlotTitle <- textGrob(paste0("Xu Data: NK vs immune cell subsets: median expression of Cand_NKsig (49 genes) \n "), gp=gpar(fontface="bold", fontsize=20))

do.call(grid.arrange, c(p, list(ncol = 3, top = PlotTitle)))
dev.off()

# the 49 genes perform pretty well except for in gammaDelta T cells (and MAIT, but still probably ok as we will have DEGs for NK vs gamDel in the DownSet). Because we took the Union, some of the poorer genes that passed in Linsley (due to there only being 6 cell types in that data set, so lower resolution in that data set cf/ the Xu data where there are ~15 cell types) creep back in

# MOYM2 is the gene that can be very highly expressed in some other blood cell subsets, but always >> greater expression in NK cells (so maybe it can be kept?)

```

NOW MOVE TO MARKDOWN #2 - previous (long) script stored in Archives under "Full Bulk RNAseq NK Biplot Sig ..."



