---
title: "Tumour and Stromal Filtration for 82-gene candidate NK sig (post scRNAseq integration"
author: "C Shembrey S Foroutan"
date: 
output:
  html_document: default
  pdf_document: default
---
In this document, we further refine our NK geneset by filtering out (i) genes that have higher expression in non-NK immune cells and (ii) genes that have high expression in tumour/epithelial cells

# Setup
```{r results = 'hide', message = FALSE, warning = FALSE}
library(limma)
library(scales)
library(ggplot2)
library(ggrepel)
library(plotly)
library(gridExtra)
library(grid)
library(ggpubr) # for statistics
library(RColorBrewer)
library(matrixStats)
library(plyr)
library(dplyr)
library(tidyr)

source("./script/biplot_ggplot.R")
```

# First, we import the candidate bulk and scRNAseq candidate genes (82 genes):
```{r}
NKUpSet_SourceTable <- read.table("../NK_candidates_Zhang_SMARTseq_10X_sc_bulk_Shembrey_Foroutan.txt", sep = "\t", header = T)
NKGenes <- alias2SymbolUsingNCBI(NKUpSet_SourceTable$Genes, "../data/Homo_sapiens.gene_info", required.columns = c("GeneID", "Symbol"))

rownames(NKGenes) <- NKGenes$GeneID
all(NKUpSet_SourceTable$Genes == NKGenes$Symbol) # T
rownames(NKUpSet_SourceTable) <- NKGenes$GeneID
```

# Next, we generate table of NK Signature sources:
```{r}
CiberAct <- read.csv("../data/NK_Signatures/Compiled/Cibersort_NK-Active.csv")
CiberAct <- NKGenes$Symbol %in% CiberAct$Genes
SumTable <-  cbind(NKGenes, CiberAct)

CiberRest <- read.csv("../data/NK_Signatures/Compiled/Cibersort_NK-Resting.csv")
Cibersort_Resting <- NKGenes$Symbol %in% CiberRest$Genes
SumTable <-  cbind(SumTable, Cibersort_Resting)

CursonsSups <- read.csv("../data/NK_Signatures/Compiled/CursonsSups_112Genes.csv")
CursonsSups <- NKGenes$Symbol %in% CursonsSups$Cursons.Supplementary.Data
SumTable <-  cbind(SumTable, CursonsSups)

MergedDEGs_Bulk <- readRDS("../data/KeyData/blood_nk/NK_genelist_2bld.RDS")
NK_upSet <- MergedDEGs_Bulk$NK_upSet
DEGs_2BldBulk <- NKGenes$Symbol %in% NK_upSet
SumTable <-  cbind(SumTable, DEGs_2BldBulk)

DEGs_NKvsPanT <- MergedDEGs_Bulk$NK_panT_up
DEGs_NKvsPanT <- NKGenes$Symbol %in% DEGs_NKvsPanT
SumTable <-  cbind(SumTable, DEGs_NKvsPanT)

RecepAct <- read.csv("../data/NK_Signatures/Compiled/NK_ReceptorActivating.csv")
RecepInhib <- read.csv("../data/NK_Signatures/Compiled/NK_ReceptorInhibitory.csv")
Receptors <- rbind(RecepAct, RecepInhib) # remove dual function receptors
Receptors <- NKGenes$Symbol %in% Receptors$Genes
SumTable <-  cbind(SumTable, Receptors)

WangGenes <- read.csv("../data/GeneSignatures/Immune_stroma/Stroma_Immune_markers_Wang_2017_Gastroenterology.csv", stringsAsFactors = F)
WangGenes <- WangGenes[WangGenes$Marker.type == "NK cell", ]
WangGenes <- WangGenes$Gene.Name
WangGenes <- NKGenes$Symbol %in% WangGenes
SumTable <-  cbind(SumTable, WangGenes)

all(rownames(SumTable) == rownames(NKUpSet_SourceTable)) # T
NKUpSet_SourceTable2 <-  as.data.frame(lapply(NKUpSet_SourceTable,FUN = as.logical))
all((NKUpSet_SourceTable2$Pass10X == "1") == (NKUpSet_SourceTable$Pass10X == "TRUE")) # T
SumTable <- c(SumTable, NKUpSet_SourceTable2)
SumTable <- as.data.frame(SumTable)

# Don't include these columns in Sum of Sources calculations:
ExclCols <- c("GeneID", "Symbol",   "Genes",  "SingleCell",   "Bulk",                  "Common_sc_bulk",  "PassSMARTseq",   "Pass10X", "Zero_95Percentile_10X")
  
SumTable$SumSources <- rowSums(SumTable[! colnames(SumTable) %in% ExclCols])
```

# First, we import bulk RNAseq data from two sources of sorted PBMCs (GSE60424 and GSE107011). We will use these to determine which genes have higher expression in NK cells relative to other immune cell types.
```{r}
# Linsley Data (GSE60424)
dataList <- readRDS("../data/NK_Signatures/CancerBlood_dataList_CRC_logRPKM.RDS") 
LinsBld <- dataList$bldCells_logRPKM
Lins <- LinsBld[rownames(LinsBld) %in% NKGenes$GeneID, ] 

LinsAnnot <- dataList$bldCells_annot 
LinsAnnot$celltype_s <- as.character(LinsAnnot$celltype_s)
LinsNKs <- LinsAnnot[LinsAnnot$celltype_s == "NK", ]
LinsNKs <- rownames(LinsNKs)
Lins_NKs <- Lins[ , colnames(Lins) %in% LinsNKs] # only the 4 NK samples
Lins_NKs <- as.data.frame(Lins_NKs)
Lins_NKs <- Lins_NKs[NKGenes$GeneID, ]
rownames(Lins_NKs) <- NKGenes$Symbol

# Xu Data
Xu <-  read.table("../data/KeyData/blood_nk/XuBulk_logTPMFilt.txt", check.names = F) # 114 samples
Xu <- Xu[rownames(Xu) %in% NKGenes$GeneID, ] 
XuAnnot <- read.csv("../data/KeyData/blood_nk/XuBulk_Annot.csv")
rownames(XuAnnot) <- XuAnnot$Samples
XuAnnot <- XuAnnot[XuAnnot$CellType == "NK", ]
Xu_NKs <- rownames(XuAnnot)
Xu_NKs <- Xu[ , colnames(Xu) %in% Xu_NKs] # only the 4 NK samples
Xu_NKs <- Xu_NKs[NKGenes$GeneID, ]
rownames(Xu_NKs) <- NKGenes$Symbol
```

# Next, we import tumour cell data from the Cancer Cell Line Encyclopedia (CCLE) and another dataset (GSE90830, Wang et al.,) containing 44 colorectal cell lines. We will use these to filter out candidate genes that have relatively high expression in tumour cells
```{r}
# CCLE Data
dataList <- readRDS("../data/NK_Signatures/CancerBlood_dataList_CRC_logRPKM.RDS") 
CCLE <- dataList$ccle_logRPKM

# CCLE   
comGenes3 <- intersect(NKGenes$GeneID, rownames(CCLE))
CCLE_NKSig <- CCLE[comGenes3, ]
dim(CCLE_NKSig) # 50 x 57 --> we miss 2 genes

missingGenes3 <- NKGenes$GeneID[!NKGenes$GeneID %in% rownames(CCLE)]
missingMat3 <- matrix(NA, nrow = length(missingGenes3), ncol = ncol(CCLE))
rownames(missingMat3) <- missingGenes3

CCLE_NKSig <- rbind(CCLE_NKSig, missingMat3)
CCLE_NKSig <- CCLE_NKSig[NKGenes$GeneID, ]
rownames(CCLE_NKSig) <- NKGenes$Symbol
# Wang Data
Wang <- dataList$wangCLs_logRPKM

comGenes4 <- intersect(NKGenes$GeneID, rownames(Wang))
Wang_NKSig <- Wang[comGenes4, ]
dim(Wang_NKSig) 
missingGenes4 <- NKGenes$GeneID[!NKGenes$GeneID %in% rownames(Wang)]
missingMat4 <- matrix(NA, nrow = length(missingGenes4), ncol = ncol(Wang))
rownames(missingMat4) <- missingGenes4

Wang_NKSig <- rbind(Wang_NKSig, missingMat4)
Wang_NKSig <- Wang_NKSig[NKGenes$GeneID, ]
rownames(Wang_NKSig) <- NKGenes$Symbol
```

```{r fig.width = 20, fig.height = 10}
# make list of tumour data
TumDataList <- list(CCLE_NKSig = CCLE_NKSig,
                    Wang_NKSig = Wang_NKSig)

TumDataList_names <- c("CCLE_NKSig",
                       "Wang_NKSig")

# get row medians for biplot point labels (later)
rowMedsTumData <- as.data.frame(lapply(TumDataList, function(x){ 
   rowMedians(as.matrix(x))
  }))

CCLE_NKSig <- as.data.frame(CCLE_NKSig)
CCLE_NKSig$RowMeds <- rowMedsTumData$CCLE_NKSig
all(rowMedians(as.matrix(CCLE_NKSig[ , -c(58)])) == CCLE_NKSig$RowMeds, na.rm = T)
CCLE_NKSig$PosRowMeds <- CCLE_NKSig$RowMeds
CCLE_NKSig$PosRowMeds <- ifelse(CCLE_NKSig$PosRowMeds >2 , rownames(CCLE_NKSig), "")
CCLE_NKSig$PosRowMeds[is.na(CCLE_NKSig$PosRowMeds)] <- ""

Wang_NKSig <- as.data.frame(Wang_NKSig)
Wang_NKSig$RowMeds <- rowMedsTumData$Wang_NKSig
all(rowMedians(as.matrix(Wang_NKSig[ , -c(45)])) == Wang_NKSig$RowMeds, na.rm = T)
Wang_NKSig$PosRowMeds <- Wang_NKSig$RowMeds
Wang_NKSig$PosRowMeds <- ifelse(Wang_NKSig$PosRowMeds >0 , rownames(Wang_NKSig), "")
Wang_NKSig$PosRowMeds[is.na(Wang_NKSig$PosRowMeds)] <- ""

# generate biplot annotations
sourceAnnot <- SumTable$SumSource
names(sourceAnnot) <- SumTable$Symbol

# Tum data vs NKdata_Lins
# The threhold should remain at 2 for the CCLE data, then reduced to 0 for the org and Wang data

# CCLE
pointAnnotCCLE <- CCLE_NKSig$PosRowMeds
names(pointAnnotCCLE) <- rownames(CCLE_NKSig)

BP1 <- biplot(
  data.expr = Lins_NKs,
  data.noexpr = TumDataList$CCLE_NKSig,
  data.names = c("NK Cells (GSE60424)", "CCLE"),
  data.expr.th = 0.5,
  data.noexpr.th = 0.5,
  annot = sourceAnnot,
  annot.name = "Sum of Sources",
  point.labels = pointAnnotCCLE,
  text.size = 0.75
  ) +
 geom_vline(xintercept = 2, col = "darkred", linetype = "dashed") +
 theme(aspect.ratio = 0.4, text = element_text(size = 22))

# Wang Cell Lines
pointAnnotWang <- Wang_NKSig$PosRowMeds
names(pointAnnotWang) <- rownames(Wang_NKSig)

BP2 <- biplot(
  data.expr = Lins_NKs,
  data.noexpr = TumDataList$Wang_NKSig,
  data.names = c("NK Cells (GSE60424)", "CRC cell lines (GSE90830)"),
  data.expr.th = 0.5,
  data.noexpr.th = 0.5,
  annot = sourceAnnot,
  annot.name = "Sum of Sources",
  point.labels = pointAnnotWang,
  text.size = 0.75
  ) +
  geom_vline(xintercept = 0, col = "darkred", linetype = "dashed") +
  theme(aspect.ratio = 0.4, text = element_text(size = 22))

PlotTitle <- textGrob(paste0("NK vs tumour cell data: median expression of biplot-filtered NK signature genes (from both bulk and scRNAseq) in Linsley Data \n Threshold = 0, keep bottom right quadrant"), gp=gpar(fontface="bold", fontsize=20))

#pdf("200708_TumFiltBiplots_Linsley_82BulkAndscRNAseqGenes.pdf", width = 30, height = 5, useDingbats = F)
grid.arrange(BP1, BP2, nrow = 1)
dev.off()

# get the genes that fail (ie. are expressed too highly in tumour data):

# Remember to remove the new RowMeds and RowMedsPos columns
data.expr.th = 0.5
data.noexpr.th = 0.5

Tum_failLins1 <- names(which(
              #apply(Lins_NKs, 1, quantile, probs = data.expr.th, na.rm = T) <= 4 & 
              apply(CCLE_NKSig[ , -c(58:59)], 1, quantile, probs = data.noexpr.th, na.rm = T) >= 2))

Tum_failLins2 <- names(which(
              #apply(Lins_NKs, 1, quantile, probs = data.expr.th, na.rm = T) <= 4 & 
              apply(Wang_NKSig[ , -c(45:46)], 1, quantile, probs = data.noexpr.th, na.rm = T) >= 0))

Final <- Reduce(intersect, list(Tum_failLins1, Tum_failLins2))
FinalUnion <- Reduce(union, list( Tum_failLins1, Tum_failLins2))

## Repeat for Xu data:

# Tum data vs NKdata_Xu
#pdf("../200501_Biplots_Candidate49Genes_TumourFiltration_Xu_withLabels.pdf", width = 25, height = 4, useDingbats = F)

# CCLE
BP1X <- biplot(
  data.expr = Xu_NKs,
  data.noexpr = TumDataList$CCLE_NKSig,
  data.names = c("NK Cells (GSE107011)", "CCLE"),
  data.expr.th = 0.5,
  data.noexpr.th = 0.5,
  annot = sourceAnnot,
  annot.name = "Sum of Sources",
  point.labels = pointAnnotCCLE,
  text.size = 0.75
  ) +
  geom_vline(xintercept = 2, col = "darkred", linetype = "dashed") +
 theme(aspect.ratio = 0.4, text = element_text(size = 22))

# Wang Cell Lines
BP2X <- biplot(
  data.expr = Xu_NKs,
  data.noexpr = TumDataList$Wang_NKSig,
  data.names = c("NK Cells (GSE107011)", "CRC cell lines (GSE90830)"),
  data.expr.th = 0.5,
  data.noexpr.th = 0.5,
  annot = sourceAnnot,
  annot.name = "Sum of Sources",
  point.labels = pointAnnotWang,
  text.size = 0.75
  ) +
  geom_vline(xintercept = 0, col = "darkred", linetype = "dashed") +
  theme(aspect.ratio = 0.4, text = element_text(size = 22))

#pdf("200708_TumFiltBiplots_Xu_82BulkAndscRNAseqGenes.pdf", width = 30, height = 5, useDingbats = F)
grid.arrange(BP1X, BP2X, nrow = 1)
dev.off()

# get the genes that fail (again, update thresholds for Org and Wang)
Tum_failXu1 <- names(which(
              #apply(Xu_NKs, 1, quantile, probs = data.expr.th, na.rm = T) <= 5 & 
              apply(CCLE_NKSig[ , -c(58:59)], 1, quantile, probs = data.noexpr.th, na.rm = T) >= 2))

Tum_failXu2 <- names(which(
              #apply(Xu_NKs, 1, quantile, probs = data.expr.th, na.rm = T) <= 5 & 
              apply(Wang_NKSig[ , -c(45:46)], 1, quantile, probs = data.noexpr.th, na.rm = T) >= 0))

Final2 <- Reduce(intersect, list(Tum_failXu1, Tum_failXu2))
FinalUnion2 <- Reduce(union, list(Tum_failXu1, Tum_failXu2))

```

# Remove the tumour-high genes 
```{r}
Union2Bld <- union(Final, Final2) 
Union2Bld_Int <- intersect(Final, Final2) 
NKGenes <- NKGenes[! NKGenes$Symbol %in% Union2Bld, ] 
```

# Next, we filter against stromal and tumour-supportive cell types (ie. other non-immune, non-tumour cells, like epithelial cells and fibroblasts) using bulk RNAseq data of these sorted populations from GSE39397 (Calon et al.,)

```{r fig.width=20, fig.height=20}
Stroma <- read.csv("../data/KeyData/CRC_Stroma/Calon8pts_CRCStroma_FilteredCounts.csv", stringsAsFactors = F)
rownames(Stroma) <- Stroma$X
Stroma <- Stroma[ , -c(1)]

# TumFilt <- read.csv("./data/NK_Signatures/Post_scRNAseq_TumStrmFIlt.csv", stringsAsFactors = F)
# TumFilt <- TumFilt[TumFilt$Fail_Source == "Fail_Tumour", ] # 12 genes
# 
# TumFilt <- NKGenes[NKGenes$Symbol %in% TumFilt$Genes, ]
# rownames(TumFilt) <- TumFilt$GeneID
# 
# Stroma_NKSigTum <- Stroma
# comGenesStr <- intersect(TumFilt$GeneID, rownames(Stroma_NKSigTum))
# Stroma_NKSigTum <- Stroma_NKSigTum[comGenesStr, ] # no missing genes
# all(rownames(Stroma_NKSigTum) == TumFilt$GeneID)


comGenesStr <- intersect(NKGenes$GeneID, rownames(Stroma))
Stroma2 <- Stroma[comGenesStr, ] 
missingGenesStr <- NKGenes$GeneID[!NKGenes$GeneID %in% rownames(Stroma)]
missingMatStr <- matrix(NA, nrow = length(missingGenesStr), ncol = ncol(head(Stroma2)))
rownames(missingMatStr) <- missingGenesStr
colnames(missingMatStr) <- colnames(Stroma2)
Stroma <- rbind(Stroma2, missingMatStr) 

StromaAnnot <- read.csv("../data/KeyData/CRC_Stroma/Calon8pts_CRCStroma_GSE39397-GPL13158_Annot.csv", stringsAsFactors = F)
# all(StromaAnnot$GEO_accession == colnames(Stroma)) --> T

# add cell type annots
Stroma <- as.data.frame(t(Stroma))
all(rownames(Stroma) == StromaAnnot$GEO_accession) # T
Stroma$CellType <- StromaAnnot$CellType

# get gene symbols
StromaReOrd <- Stroma[ , c(NKGenes$GeneID, "CellType")]
all(colnames(StromaReOrd)[1:69] == NKGenes$GeneID) # T
colnames(StromaReOrd)[1:69] <- NKGenes$Symbol
StromaReOrd$Sample <- rownames(StromaReOrd)

library(reshape2)
meltTest <- melt(StromaReOrd, id.vars = c("CellType", "Sample"), na.rm = T)
colnames(meltTest)[3:4] <- c("Gene", "Expression")
meltTest$CellType <- factor(meltTest$CellType, levels = c("Leukocytes", "Stromal cells", "Epithelial cells", "Fibroblasts"))

# check the ordering
#meltTest2 <- meltTest[order(meltTest$CellType), ]
#meltTest <- meltTest[complete.cases(meltTest$Gene), ]

Alphabetise <- as.character(unique(meltTest$Gene))
Alphabetise <- sort(Alphabetise)
meltTest$Gene = factor(meltTest$Gene, levels=Alphabetise)

# meltTest_Split <- split(meltTest, f = as.factor(meltTest$Gene), drop = F)
# meltTest_Split_names <- levels(meltTest$Gene)
# names(meltTest_Split) <- meltTest_Split_names
```

# Generate Figure S3
```{r}
# for stats:
SigLvls <- list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns"))

#pdf("Post_SingleCell_StromalFilt.pdf", useDingbats = F)
my_comparisons <- list(c("Leukocytes", "Stromal cells"), c("Leukocytes", "Fibroblasts"),c("Leukocytes", "Epithelial cells"))

#pdf("200708_Faceted_StromalFiltBoxplots.pdf", width = 20, height = 30)
ggplot(meltTest, aes(x = CellType, y = Expression, fill = CellType, color = CellType)) + 
        geom_boxplot(alpha = 0.6, outlier.colour = NA, color = "black") +
        geom_jitter(width = 0.1, size = 3) +
        ggtitle("Stromal Filtration (82 candidate genes; bulk + scRNAseq candidates)") +
        labs(x= NA, y = "Expression (logRPKM)") +
        scale_fill_manual(values = c("purple", "goldenrod1", "springgreen4", "cornflowerblue")) +
        scale_color_manual(values = c("purple", "darkgoldenrod", "springgreen4", "cornflowerblue")) +
        stat_compare_means(comparisons = my_comparisons, label = "p.signif", size = 5) +
        # scale_y_continuous(expand = expand_scale(mult = c(0, 0.15))) +
        scale_y_continuous(expand = expand_scale(mult = 0.15)) +
        facet_wrap(~Gene, scales = "free_y", ncol = 7) +
        theme_light() +
        theme(legend.position = "none",
               axis.title = element_blank(),
               axis.text.x = element_blank(),
               plot.title = element_text(hjust = 0.5, size = 22),
               #aspect.ratio = 1,
               text = element_text(size = 14),
               panel.grid.minor = element_blank(),
               strip.background = element_rect(fill = "black")
              )
dev.off()
```

```{r}
TestMeans <- meltTest
TestMeans$Gene_CellType <- paste(TestMeans$Gene, TestMeans$CellType, sep = "_")
TestMeans$Gene_CellType <- as.factor(TestMeans$Gene_CellType)

TestMeans <- ddply(TestMeans, .(Gene_CellType), summarise, mean=mean(Expression)) # 1736 rows
TestMeans$ToSplit <- TestMeans$Gene_CellType
TestMeans <- separate(data = TestMeans, col="ToSplit", into = c("Gene", "CellType"), sep = "_")
TestMeans$Gene <- as.factor(TestMeans$Gene)
TestMeans$CellType <- as.factor(TestMeans$CellType)
colnames(TestMeans)[2] <- "MeanExpr"

ByGene <- split(TestMeans, TestMeans$Gene)


ByGene2 <- lapply(ByGene, function(x){
  x %>% mutate(NormalisedToNK = MeanExpr / MeanExpr[CellType == "Leukocytes"])})
TestMeans <- bind_rows(ByGene2, .id = "")
TestMeans <- TestMeans[ , -c(1)]
```

# Generate figure 2C
```{r fig.width=20, fig.height=8}
TestMeans$CellType <- factor(TestMeans$CellType, levels = c("Leukocytes", "Stromal cells", "Fibroblasts", "Epithelial cells"))
TestMeans <- TestMeans[order(TestMeans$CellType), ]
TestMeans$Gene <- factor(TestMeans$Gene, levels = unique(TestMeans$Gene))

pdf("StromalFilt_BoxPlotSummary_Linegraph.pdf", height = 7, width = 20, useDingbats = F)
ggplot(TestMeans, 
       aes_string(x="Gene", y = "NormalisedToNK", color = "CellType", group = "CellType")) + 
  geom_point(size = 5) +
  ylab("Expression (Normalised to Leukocytes)") +
  geom_path() +
  scale_color_manual(values = c("purple", "goldenrod", "springgreen4", "cornflowerblue")) +
  theme_light() +
  theme(text = element_text(size = 20), 
        axis.text.x = element_text(angle = 90))
dev.off()
```
