---
title: "Kaplan-Meier Curves - NK Score"
author: "C Shembrey S Foroutan"
date: "19/11/2020"
output: html_document
---
In this document, we take the NK scores (generated in previous markdown) for the TCGA and Marisa et al., datasets and perform survival analysis to determine whether high NK score is associated with any survival outcomes.

# First, we read in a csv file where survival outcomes for the TCGA patients and their NK scores have been merged, as well 
```{r}
library(survival)
library(survminer)
library(ggplot2)

TCGABind <- read.csv("./data/TCGABind_ForSurvivalAnalysis.csv", stringsAsFactors = F, row.names = "X") 

# read in Clinical annot
ClinAnnot <- read.csv("../data/KeyData/TCGA/TCGA-COAD_ClinicalReponseAnnot.csv", stringsAsFactors = F)
ClinAnnot$bcr_patient_barcode <-  gsub("-", ".", ClinAnnot$bcr_patient_barcode)
summary(rownames(TCGABind) %in% ClinAnnot$bcr_patient_barcode) # 454/454 true
rownames(ClinAnnot) <- ClinAnnot$bcr_patient_barcode
ClinAnnot <- ClinAnnot[rownames(TCGABind), ] # shrink to 454 common samples
TCGABind <- cbind(TCGABind, ClinAnnot[ , c(3,4,25:32)])
TCGABind[ , c(29:36)] <- lapply(TCGABind[ , c(29:36)], as.numeric) # time variables

# We remove the Stage IV patients, as metastatic disease should not be grouped together with primary disease for survival analysis:
TCGABind_noIV <- TCGABind[!TCGABind$tumor_stage == "Stage IV", ]

# group the patients according to whether their NK score is above or below the median NK score for the cohort
median(TCGABind_noIV$NKUp, na.rm = T) # -0.3944445
TCGABind_noIV$NKScore_SplitMed <- TCGABind_noIV$NKUp
TCGABind_noIV$NKScore_SplitMed <- ifelse(TCGABind_noIV$NKScore_SplitMed > -0.394445, "NK.High", "NK.Low")

```

# Genmerate Kaplan-Meier curves for NK High vs Low sample groupings in TCGA data
```{r fig.width=20}

# TCGA NK Score
  # median(TCGABind$NKUp, na.rm = T) # -0.3956979
  # TCGABind$NKScore_SplitMed <- TCGABind$NKUp
  # TCGABind$NKScore_SplitMed <- ifelse(TCGABind$NKScore_SplitMed > -0.3956979, "NK.High", "NK.Low")

# Unstaged
TCGAfit_OS <- survfit(Surv(OS.time, OS)~NKScore_SplitMed , data = TCGABind_noIV)
TCGAfit_PFI <- survfit(Surv(PFI.time, PFI)~NKScore_SplitMed , data = TCGABind_noIV)
TCGAfit_DFI <- survfit(Surv(DFI.time, DFI)~NKScore_SplitMed , data = TCGABind_noIV[complete.cases(TCGABind_noIV$DFI), ])

# FIGURES FOR THESIS
#pdf("Kaplan-Meier_Curves_TCGA_NKScore.pdf", width = 10, height = 8)
ggsurvplot(TCGAfit_OS, data = TCGABind_noIV,
                  #facet.by = "tumor_stage",
           conf.int = F,
           pval = T,
           risk.table = T,
           palette = c("darkred", "blue")) +  ggtitle("Shembrey NK Score / TCGA-COAD (OS) \n NK-High = above median NK score, NK-Low = below median NK Score")

ggsurvplot(TCGAfit_PFI, data = TCGABind_noIV,
                  #facet.by = "tumor_stage",
           conf.int = F,
           pval = T,
           risk.table = T,
           palette = c("darkred", "blue")) +  ggtitle("Shembrey NK Score / TCGA-COAD (PFI) \n NK-High = above median NK score, NK-Low = below median NK Score")

ggsurvplot(TCGAfit_DFI, data = TCGABind_noIV,
                  #facet.by = "tumor_stage",
           conf.int = F,
           pval = T,
           risk.table = T, 
           palette = c("darkred", "blue")) +  ggtitle("Shembrey NK Score / TCGA-COAD (DFI) \n NK-High = above median NK score, NK-Low = below median NK Score")
dev.off()
```

# Calculate significance using Lasso (for TCGA):
```{r}
## Define survival parameters:
timeCol <- "DFI.time"
eventCol <- "DFI"

## subset your annotation data to columns you want to use (all clinical factors of interests as well as survivlal outcome)
## I am also removing all the normal samples
Annot_ForLasso <-
  TCGABind_noIV[ , c("age_at_initial_pathologic_diagnosis",
                       "tumor_stage",
                       "Subtype_MSI_Status",
                        "NKScore_SplitMed",
                       timeCol,
                       eventCol
                     )]

## make sure you don't have NA in the data and that you survival time is > 0. 
Annot_ForLasso <- Annot_ForLasso[complete.cases(Annot_ForLasso) &
                                             Annot_ForLasso[, timeCol] > 0,]

## get columns only for the predictors (clinical variables) by removing survival columns
pred <- Annot_ForLasso[,!colnames(Annot_ForLasso) %in% c(timeCol, eventCol)]
## make the class of the predictors appropriate for the glmnet function; 
## you need to add the column names of all the clinical factors here
x <- model.matrix(~ age_at_initial_pathologic_diagnosis + tumor_stage + Subtype_MSI_Status, pred)

## define output data, this is survival data and has to have the below column names
y <- as.matrix(Annot_ForLasso[, c(timeCol, eventCol)])
colnames(y) <- c("time", "status")

## run glmnet for lasso, we then look at the plot
library(glmnet)
fit <- glmnet(
  x = x, 
  y = y, 
  family = "cox",
  alpha = 1)

plot(fit, label = T)

## perform cross validation for to obtain coef for the variables at min lambda
cv.fit <- cv.glmnet(x, 
                    y, 
                    family = "cox", 
                    alpha = 1)
plot(cv.fit)

## We need to include the variables in the Cox model, if we have non-zero (non-dot) for any of the elements of a given variable.
coef(cv.fit, s = "lambda.min")

## Then run coxph by only including the variables significant in the above test as well as your NK signature: 
coxph(
  Surv(
    as.numeric(Annot_ForLasso[, timeCol]),
    as.numeric(Annot_ForLasso[, eventCol])
  ) ~
    age_at_initial_pathologic_diagnosis + tumor_stage + Subtype_MSI_Status + NKScore_SplitMed,
  data = Annot_ForLasso
)

```
# Generate Kaplan-Meier curves for Marisa Data
```{r fig.height = 8, fig.width = 10}
Marisa <- read.csv("./data/Marisa_ForSurvivalAnalysis.csv", row.names = "X", stringsAsFactors = F) # there is RFS event/delay and OS event/delay IN MONTHS
Marisa[ , c(14:17)] <- lapply(Marisa[ , c(14:17)], as.numeric) # time variables
Marisa <- Marisa[! Marisa$TNM_Stage == "0", ]

# Marisa NK Score
    # median(Marisa$NKUp, na.rm = T) # -0.3853625
    # Marisa$NKScore_SplitMed <- Marisa$NKUp
    # Marisa$NKScore_SplitMed <- ifelse(Marisa$NKScore_SplitMed > -0.3853625, "NK.High", "NK.Low")

# remove stage IV samples with RFS = 0
tt <- Marisa[Marisa$TNM_Stage == "4", ] # subset to stage IV
tt <- tt[tt$RFS_Delay == "0", ]
Marisa_noIV <- Marisa[! rownames(Marisa) %in% rownames(tt), ]

median(Marisa_noIV$NKUp, na.rm = T) # -0.38494
Marisa_noIV$NKScore_SplitMed <- Marisa_noIV$NKUp
Marisa_noIV$NKScore_SplitMed <- ifelse(Marisa_noIV$NKScore_SplitMed > -0.38494, "NK.High", "NK.Low")

# Unstaged
MarisaFit_OS <- survfit(Surv(OS_Delay, OS_Event)~NKScore_SplitMed , data = Marisa_noIV) #+ TNM_Stage
MarisaFit_RFS <- survfit(Surv(RFS_Delay, RFS_Event)~NKScore_SplitMed , data = Marisa_noIV) #+ TNM_Stage

MarisaFit_RFS2 <- survfit(Surv(RFS_Delay, RFS_Event)~NKScore_SplitMed , data = Marisa_noIV) #+ TNM_Stage

#pdf("Marisa_OS_RFS_ForThesis.pdf", width = 10, height=10) # need to remove Stage IV patients
ggsurvplot(MarisaFit_RFS2, data = Marisa_noIV,
                  #facet.by = "tumor_stage",
           conf.int = F,
           pval = T,
           palette = c("darkred", "blue"),
           risk.table = T) +  
           ggtitle("NK Score / Marisa (RFS) \n NK-High = above median NK score, NK-Low = below median NK Score")

ggsurvplot(MarisaFit_OS, data = Marisa_noIV,
                  #facet.by = "tumor_stage",
           conf.int = F,
           pval = T,
           palette = c("darkred", "blue"),
           risk.table = T) +  
           ggtitle("NK Score / Marisa (OS) \n NK-High = above median NK score, NK-Low = below median NK Score")
dev.off()

```

# Calculate Cox P-values, controlling for age and tumour stage and MSI status
```{r}
TCGABind_noIV$age_at_initial_pathologic_diagnosis <- as.numeric(TCGABind_noIV$age_at_initial_pathologic_diagnosis)

median(TCGABind_noIV$age_at_initial_pathologic_diagnosis, na.rm = T) # 69

TCGABind_noIV$BinaryAge <- TCGABind_noIV$age_at_initial_pathologic_diagnosis
TCGABind_noIV$BinaryAge <- ifelse(TCGABind_noIV$BinaryAge > 69, "Old", "Young")
TCGABind_noIV$BinaryAge <- as.factor(TCGABind_noIV$BinaryAge)

TCGABind_noIV$BinaryMSI <- TCGABind_noIV$Subtype_MSI_Status
TCGABind_noIV$BinaryMSI <- ifelse(TCGABind_noIV$BinaryMSI == "MSI-H", "MSI", "MSS")
TCGABind_noIV$BinaryMSI <- as.factor(TCGABind_noIV$BinaryMSI)

res.coxTCGA <- coxph(Surv(DFI.time, DFI) ~ NKScore_SplitMed + tumor_stage +  BinaryAge + BinaryMSI, data =  TCGABind_noIV) # BinaryMSI #gender
summary(res.coxTCGA)

res.coxTCGA <- coxph(Surv(PFI.time, PFI) ~ NKScore_SplitMed + tumor_stage +  BinaryAge + BinaryMSI, data =  TCGABind_noIV) # BinaryMSI #gender
summary(res.coxTCGA)

res.coxTCGA <- coxph(Surv(OS.time, OS) ~ NKScore_SplitMed + tumor_stage +  BinaryAge + BinaryMSI, data =  TCGABind_noIV) # BinaryMSI #gender
summary(res.coxTCGA)

res.coxTCGA <- coxph(Surv(DFI.time, DFI) ~ NKScore_SplitMed + tumor_stage +  BinaryAge + BinaryMSI, data =  TCGABind_noIV) # BinaryMSI #gender
summary(res.coxTCGA)

res.coxTCGA <- coxph(Surv(PFI.time, PFI) ~ NKScore_SplitMed + tumor_stage +  BinaryAge + BinaryMSI, data =  TCGABind_noIV) # BinaryMSI #gender
summary(res.coxTCGA)

Marisa_noIV$AgeAtDiagnosis <- as.numeric(Marisa_noIV$AgeAtDiagnosis)
median(Marisa_noIV$AgeAtDiagnosis, na.rm = T) # 69
Marisa_noIV$BinaryAge <- Marisa_noIV$AgeAtDiagnosis
Marisa_noIV$BinaryAge <- ifelse(Marisa_noIV$BinaryAge > 69, "Old", "Young")
Marisa_noIV$BinaryAge <- as.factor(Marisa_noIV$BinaryAge)

Marisa_noIV$MMR_Status[Marisa_noIV$MMR_Status == "N/A"] <- NA
Marisa_noIV$MMR_Status <- as.factor(Marisa_noIV$MMR_Status)
res.coxMarisa <- coxph(Surv(RFS_Delay, RFS_Event) ~ NKScore_SplitMed + TNM_Stage + BinaryAge + MMR_Status, data =  Marisa_noIV)
summary(res.coxMarisa)

res.coxMarisa <- coxph(Surv(OS_Delay, OS_Event) ~ NKScore_SplitMed + TNM_Stage + BinaryAge + MMR_Status, data =  Marisa_noIV)
summary(res.coxMarisa)

table(Marisa_noIV$MMR_Status)
```